\chapter{Acknowledgements}

This thesis would never existed without the support and assistance of numerous people. In fact it wasn't even planned to be 
finished at all after an excessive experimental verification of Murphy's Law for a good chunk of time. Strictly speaking, Maarten Steinbuch 
persuaded me with his irresistible negotiation skills in about fifteen minutes for finishing the thesis during a casual chat that took 
place in a colloquium (for the record, there were drinks involved too). I owe him much more than this manuscript for his generous support 
and pragmatic approach to idealism (and making it look like not a contradiction). There are not many of his kind in the academic circles 
who can tell the difference between a proof and a solution.

Siep Weiland is another individual who utterly confused me about my dislike about academia (not that he is aware of it). The very few
hour-long discussions we had after becoming the co-promoter just a few months before the defense ceremony was almost an emotional 
experience. I had discussed year worth content with him in the order of minutes as if we were working together the whole time. It saddens 
me that we had met this late but I guess it's one of those stochastic jokes of life. Nevertheless I still feel lucky to have him on board 
and I thank him for everything. 

I would like to thank also to my thesis committee members especially for daring to get to the bottom of this rather unconventional text. 

I was supposed to finalize this thesis under DCSC, \textsf{\textbf{T\textcolor{cyan}{U}}Delft} affiliation. But it did not quite happen that 
way. Still I would like to thank Carsten W. Scherer for giving me the chance to work in this project and his guidance and insights during 
his time in DCSC. 

Now the difficult part; I don't know how to sum up the list of people whom I have become friends with, enjoyed their company and benefited 
from their support. After scratching many versions, because I really suck at this thing, this is what I decided to do so here goes nothing; 
Nazl\i\space who always managed to find a way to make me do things right with her love and unparalleled optimism towards life, the people 
in DCSC (I wish the best for all of you if you know what I mean), the people in CST for your spirit and comfortable environment, C\'{e}sar 
L\'{o}pez for helping me taming the experimental setup and for our technical discussions, the best secretariat in the world ever: my dearest 
Kitty, Esther, Saskia, Ellen in DCSC and Geertje and Petra in CST, the Delft Haptics Lab people for the fun, my colleagues Aram and Anton 
with whom I had to play the thesis supervisor role while they did their magic, dudes of the \enquote{pimientos} 
gang for the beerwatch nights (often, judging by its residents, in the most unfortunately named \enquote{Tango}), Rotterdam tayfas\i, Delft 
tayfas\i, my awesome band Bike2Rome, my inmates from Koornmarkt 42 for all the occasions of which, for \enquote{some} reason, I cannot 
remember much (how did a full-size table-tennis table end up in my room again?), my veteran DTXpress III whereever you are now, my 
colleagues at ASML and TMC Mechatronics, Joris for the daunting task of the translation of the summary of this thesis, all the nice people 
in \href{http://tex.stackexchange.com}{\TeX-SX} for the procrastination, thank you very much indeed!


It goes without saying that I am grateful to mom and dad for being who they are. I am still amazed how they can contain their patience and 
understanding no matter how crazy stuff I come up with; \enquote{\emph{Hocam, sizin o\u{g}lan neden b\"{o}yle?}}